NN_Accel_Heading

### 2024-03-11 Fish Data Analysis/NN_Accel_Heading.qmd
Fully flipped fish pairs to do both

fish_comp_data_switch_fish <- fish_comp_data %>% mutate(Fish3 = Fish1,
                                              Fish3_X = Fish1_X,
                                              Fish3_Y = Fish1_Y,
                                              Fish3_Z = Fish1_Z,
                                              Fish3_Speed = Fish1_Speed,
                                              Fish3_Accel = Fish1_Accel,
                                              Fish3_Yaw_Heading = Fish1_Yaw_Heading,
                                              Fish3_Yaw_Heading_PS = Fish1_Yaw_Heading_PS
                                              ) %>%
                                       mutate(Fish1 = Fish2,
                                              Fish1_X = Fish2_X,
                                              Fish1_Y = Fish2_Y,
                                              Fish1_Z = Fish2_Z,
                                              Fish1_Speed = Fish2_Speed,
                                              Fish1_Accel = Fish2_Accel,
                                              Fish1_Yaw_Heading = Fish2_Yaw_Heading,
                                              Fish1_Yaw_Heading_PS = Fish2_Yaw_Heading_PS) %>%
                                       mutate(Fish2 = Fish3,
                                              Fish2_X = Fish3_X,
                                              Fish2_Y = Fish3_Y,
                                              Fish2_Z = Fish3_Z,
                                              Fish2_Speed = Fish3_Speed,
                                              Fish2_Accel = Fish3_Accel,
                                              Fish2_Yaw_Heading = Fish3_Yaw_Heading,
                                              Fish2_Yaw_Heading_PS = Fish3_Yaw_Heading_PS) %>%
                                       #Switch around angle when we switch focal fish
                                       mutate(Angle = ifelse(Angle >= 0, Angle - pi, Angle + pi),
                                              Relative_X = -1*Relative_X) %>%
                                       select(-c(Fish3,Fish3_X,Fish3_Y,Fish3_Z,Fish3_Speed,Fish3_Accel,Fish3_Yaw_Heading,Fish3_Yaw_Heading_PS))

@@

fish_data_nn <- fish_comp_data_combo %>% group_by(Fish1,Tailbeat_Num,Data_ID) %>%
                                    mutate(#nndDistance = min(Distance, na.rm = TRUE),
                                           nndDistance = nth_min(Distance, 1),
                                           nndDistance = ifelse(is.infinite(nndDistance), NA, nndDistance),
                                           all_nas = all(is.na(nndDistance))) %>%
                                    filter(Distance == nndDistance | all_nas) %>%
                                    mutate(row_n = row_number()) %>%
                                    filter(row_n == 1) %>%
                                    ungroup() %>%
                                    mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS) %>%
                                    filter(Distance <= 1.5)

@@

nn_heading_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Angle,pi/5)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

@@

fish_data_nn_model <- fish_data_nn %>% filter(Flow == "Still Water") %>% 
                                       mutate(AblationCat = ifelse(Ablation == "Ablated",1,0),
                                              DarknessCat = ifelse(Darkness == "Dark",1,0))

@@

heading_ps_angle_step <- stepAIC(heading_ps_angle_base,
                                   list(upper = ~ (sin(Angle)+Ablation:sin(Angle)+Darkness:sin(Angle)), lower = ~1),
                                   direction = "both",
                                   trace = F)

@@

#Raw data
ggplot(fish_data_nn %>% filter(Flow == "Still Water"),
       aes(x = Angle, y = Fish1_Yaw_Heading_PS, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_line(aes(y = predict(heading_ps_angle_step)), size = 1)+
  theme_classic() +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")

#Means graph
ggplot(nn_heading_sum %>% filter(Flow == "Still Water"),
       aes(x = angle_bins, y = rad2deg(mean_turn), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  geom_line(aes(y = rad2deg(predict(manual_model,
                    newdata = nn_heading_sum %>% filter(Flow == "Still Water") %>% mutate(Angle = angle_bins)))), size = 1)+
  theme_classic() +
  ylim(-50,50) +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (degrees/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")

@@

@@

accel_base_lm <- lm(Fish1_Accel ~ Flow_Ablation_Darkness, data = fish_data_nn %>% filter(Flow == "Still Water"))
accel_lin_lm <- lm(Fish1_Accel ~ Relative_X*Flow_Ablation_Darkness, data = fish_data_nn %>% filter(Flow == "Still Water"))

@@

ggplot(fish_data_nn %>% filter(Flow == "Still Water"),
       aes(x = Relative_X, y = Fish1_Accel, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_smooth(se = F, method = glm, formula = y ~ x)+
  theme_classic() +
  xlab("distance from neighbour (BL)") +
  ylab("focal fish acceleration (BL/s^2)") +
  #ylim(-0.01, 0.01) +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  labs(color = "Sensory Conditions")

ggplot(nn_accel_sum %>% filter(Flow == "Still Water" & distance_bins < 3),
       aes(x = distance_bins, y = mean_accel, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  #geom_line(alpha = 0.75) +
  geom_smooth(se = F, method = glm, formula = y ~ 1)+
  theme_classic() +
  xlab("distance from neighbour (BL)") +
  ylab("focal fish acceleration (BL/s^2)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  labs(color = "Sensory Conditions")

@@

directional_df <- fish_data_nn %>% mutate(round_X = round_any(RelativeX,0.5),
                                                    round_Y = round_any(RelativeY,0.5)) %>%
                                                      group_by(round_X,round_Y,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                      summarise(mean_accel = mean(Fish1_Accel),
                                                                mean_turn = mean(Fish1_Yaw_Heading_PS),
                                                                n = n()) %>%
                                                      ungroup()

ggplot(directional_df %>% filter(Flow == "Still Water" & n > 5 & mean_accel > -0.01 & mean_accel < 0.01),
       aes(round_X, round_Y, fill= mean_accel)) + 
  geom_tile()+
  facet_wrap(~Flow_Ablation_Darkness)+
  scale_fill_viridis(option="magma")+
  coord_fixed()+
  theme_classic()

ggplot(directional_df %>% filter(Flow == "Still Water" & n > 5),
       aes(round_X, round_Y, fill= mean_turn)) + 
  geom_tile()+
  facet_wrap(~Flow_Ablation_Darkness)+
  scale_fill_viridis(option="magma")+
  coord_fixed()+
  theme_classic()

>>>>

fish_comp_data_switch_fish <- fish_comp_data %>% mutate(Fish3 = Fish1,
                                              Fish3_X = Fish1_X,
                                              Fish3_Y = Fish1_Y,
                                              Fish3_Z = Fish1_Z,
                                              Fish3_Speed = Fish1_Speed,
                                              Fish3_Accel = Fish1_Accel,
                                              Fish3_Yaw_Heading = Fish1_Yaw_Heading,
                                              Fish3_Yaw_Heading_PS = Fish1_Yaw_Heading_PS
                                              ) %>%
                                       mutate(Fish1 = Fish2,
                                              Fish1_X = Fish2_X,
                                              Fish1_Y = Fish2_Y,
                                              Fish1_Z = Fish2_Z,
                                              Fish1_Speed = Fish2_Speed,
                                              Fish1_Accel = Fish2_Accel,
                                              Fish1_Yaw_Heading = Fish2_Yaw_Heading,
                                              Fish1_Yaw_Heading_PS = Fish2_Yaw_Heading_PS) %>%
                                       mutate(Fish2 = Fish3,
                                              Fish2_X = Fish3_X,
                                              Fish2_Y = Fish3_Y,
                                              Fish2_Z = Fish3_Z,
                                              Fish2_Speed = Fish3_Speed,
                                              Fish2_Accel = Fish3_Accel,
                                              Fish2_Yaw_Heading = Fish3_Yaw_Heading,
                                              Fish2_Yaw_Heading_PS = Fish3_Yaw_Heading_PS) %>%
                                       #Switch around angle when we switch focal fish
                                       mutate(Angle = -1*ifelse(Angle >= 0, Angle - pi, Angle + pi),
                                              Relative_X = -1*Relative_X) %>%
                                       select(-c(Fish3,Fish3_X,Fish3_Y,Fish3_Z,Fish3_Speed,Fish3_Accel,Fish3_Yaw_Heading,Fish3_Yaw_Heading_PS))

@@

min_dist <- 0.75
max_dist <- 10

fish_data_nn <- fish_comp_data_combo %>% group_by(Fish1,Tailbeat_Num,Data_ID) %>%
                                    mutate(#nndDistance = min(Distance, na.rm = TRUE),
                                           nndDistance = nth_min(Distance, 1),
                                           nndDistance = ifelse(is.infinite(nndDistance), NA, nndDistance),
                                           all_nas = all(is.na(nndDistance))) %>%
                                    #filter(Distance == nndDistance | all_nas) %>%
                                    mutate(row_n = row_number()) %>%
                                    #filter(row_n == 1) %>%
                                    ungroup() %>%
                                    mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS) %>%
                                    filter(Distance >= min_dist & Distance <= max_dist)

@@

nn_heading_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Angle,pi/6)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

nn_heading_dist_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Angle,pi/6), distance_bins = round_any(Distance,0.25)) %>%
                                                group_by(angle_bins,distance_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

@@

fish_data_nn_model <- fish_data_nn_normal %>% filter(Flow == "Still Water") %>% 
                                       mutate(AblationCat = ifelse(Ablation == "Ablated",1,0),
                                              DarknessCat = ifelse(Darkness == "Dark",1,0))

@@

heading_ps_angle_step <- stepAIC(heading_ps_angle_base,
                                   list(upper = ~ Angle*sin(Angle)*Ablation*Darkness, lower = ~1),
                                   direction = "both",
                                   trace = F)

@@

manual_model_2 <- lm(Fish2_Yaw_Heading_PS ~ sin(Angle) + sin(Angle):Ablation + sin(Angle):Darkness + sin(Angle):Ablation:Darkness, 
                   data = fish_data_nn_model %>% mutate(Angle = ifelse(Angle >= 0, Angle - pi, Angle + pi))) 
Anova(manual_model_2)

#Raw data
ggplot(fish_data_nn %>% filter(Flow == "Still Water"),
       aes(x = Angle, y = rad2deg(Fish2_Yaw_Heading_PS), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_line(aes(y = predict(heading_ps_angle_step)), size = 1)+
  theme_classic() +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")

#Means graph
ggplot(nn_heading_sum %>% filter(Flow == "Still Water"),
       aes(x = angle_bins, y = rad2deg(mean_turn), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  geom_line()+
  # geom_line(aes(y = rad2deg(predict(manual_model,
  #                   newdata = nn_heading_sum %>% filter(Flow == "Still Water") %>% mutate(Angle = angle_bins)))), size = 1)+
  geom_line(aes(y = rad2deg(predict(manual_model,
                    newdata = nn_heading_sum %>% filter(Flow == "Still Water") %>% mutate(Angle = angle_bins)))), size = 1)+
  theme_classic() +
  ylim(-50,50) +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (degrees/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")+
  ggtitle(paste0("All Fish From ",min_dist," to ",max_dist))

@@

# ggplot(nn_heading_dist_sum %>% filter(Flow == "Still Water"),
#        aes(x = angle_bins, y = distance_bins, fill = mean_turn))+
#   facet_wrap(~ Ablation + Darkness)+
#   scale_fill_viridis(option="turbo",limits = c(-2,2))+
#   geom_tile()+
#   scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
#                      labels = round(unique(nn_heading_sum$angle_bins))) +
#   coord_polar(start = -pi / 2)+
#   theme_classic()

@@

accel_base_lm <- lm(Fish1_Accel ~ Flow_Ablation_Darkness, data = fish_data_nn %>% filter(Flow == "Flowing Water (2 BL/s)"))
accel_lin_lm <- lm(Fish1_Accel ~ Relative_X*Flow_Ablation_Darkness, data = fish_data_nn %>% filter(Flow == "Flowing Water (2 BL/s)"))

@@

ggplot(fish_data_nn %>% filter(Flow == "Flowing Water (2 BL/s)"),
       aes(x = Relative_X, y = Fish1_Accel, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  geom_smooth(se = F, method = glm, formula = y ~ x)+
  theme_classic() +
  xlab("distance from neighbour (BL)") +
  ylab("focal fish acceleration (BL/s^2)") +
  #ylim(-0.01, 0.01) +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  labs(color = "Sensory Conditions")

ggplot(nn_accel_sum %>% filter(Flow == "Flowing Water (2 BL/s)" & distance_bins < 3),
       aes(x = distance_bins, y = mean_accel, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  #geom_line(alpha = 0.75) +
  geom_smooth(se = F, method = glm, formula = y ~ x)+
  theme_classic() +
  xlab("distance from neighbour (BL)") +
  ylab("focal fish acceleration (BL/s^2)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  labs(color = "Sensory Conditions")

@@

# directional_df <- fish_data_nn %>% mutate(round_X = round_any(Relative_X,0.5),
#                                                     round_Y = round_any(Relative_Y,0.5)) %>%
#                                                       group_by(round_X,round_Y,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
#                                                       summarise(mean_accel = mean(Fish1_Accel),
#                                                                 mean_turn = mean(Fish1_Yaw_Heading_PS),
#                                                                 n = n()) %>%
#                                                       ungroup()
# 
# ggplot(directional_df %>% filter(Flow == "Still Water" & n > 5 & mean_accel > -0.01 & mean_accel < 0.01),
#        aes(round_X, round_Y, fill= mean_accel)) + 
#   geom_tile()+
#   facet_wrap(~Flow_Ablation_Darkness)+
#   scale_fill_viridis(option="magma")+
#   coord_fixed()+
#   theme_classic()
# 
# 
# ggplot(directional_df %>% filter(Flow == "Still Water" & n > 5),
#        aes(round_X, round_Y, fill= mean_turn)) + 
#   geom_tile()+
#   facet_wrap(~Flow_Ablation_Darkness)+
#   scale_fill_viridis(option="magma")+
#   coord_fixed()+
#   theme_classic()
